---
title: "Test 2 Review"
author: "Matthew Swanson"
date: "10/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Use the `Recidivism` data from the `resampledata` library for this problem. Test the hypothesis that: *The mean days to `Days` to reoffend for men is higher than for women?*
i. Assume that $\mu_M$ and $\mu_W$ be the mean `Days` to reoffend for men and women respectively. Write $H_0$ and $H_a$ for this problem.
    
* $H_0: \mu_M - \mu_W = 0$
* $H_a: \mu_M - \mu_W > 0$

*Note: use* `set.seed(10001)`

ii. Conduct the test.

```{r}
library(resampledata)
data("Recidivism")
library(dplyr)
library(knitr)

ans = Recidivism %>%
  group_by(Gender) %>%
  summarize(mean = mean(Days, na.rm = TRUE), N = n())
ans
md = ans[2,2] - ans[1,2]
observed = md$mean
observed

set.seed(10001)
sims = 10^4-1 # Number of simulations (iterations)
ans = numeric(sims)
for(i in 1:sims){
  index = sample.int(n = 17019, size = 14918, replace = FALSE) #n = total, size = num in the first group
  ans[i] = mean(Recidivism$Days[index], na.rm=TRUE) - mean(Recidivism$Days[-index], na.rm=TRUE)
}
pvalue = (sum(ans >= observed) + 1)/(sims + 1)
pvalue
```

Make conclusion assuming $\alpha = 0.05$

a. Technical conclusion

```{r}
#code
```

b. Non-technical conclusion

```{r}
#code
```

