---
title: "Exam 1 Reference Sheet"
author: "Matthew Swanson"
date: "9/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(resampledata)
library(dplyr)
library(knitr)
library(nycflights13)   #get the data
```

#Vectors and Matrices
```{r}
seq_vec = seq(1,19,2)                               #seq([first],[last],[inc])
seq_mat = matrix(seq_vec, nrow = 2, byrow = TRUE)   #matrix([vec],nrow=[int],ncol=[int],byrow=[bool])
seq_mat
```

```{r include=FALSE}
nv <- c(1, 3, 6, 8)                 #make the vecs
cv <- c("a", "v", "f", "p")
lv <- c(TRUE, FALSE, FALSE, TRUE)
DF1 <- data.frame(nv, cv, lv)       #create the data frame using the vecs as args
```

```{r}
#subsets (two ways)
subset(DF1, subset = lv == TRUE)    #subset([df], subset = [lexp])
DF1[lv == TRUE, ]
subset(DF1, subset = nv > 5)        #grabs rows 3 and 4
#order
order(seq_vec)                      #ret pos of elements that'd make the ordered vector
#frame heads
#head(ChickWeight[order(ChickWeight$weight),], n = 5)    #get 5 lightest chickens
```

#List
```{r}
#my_matrix <- matrix(1:9, byrow = TRUE, nrow = 3)
#my_df <- DF1
#my_list <- list(my_vector, my_matrix, my_df)                #unnamed list
#my_list2 <- list(Vector = my_vector, Matrix = my_matrix)    #named list
#my_list2$Matrix         #selecting the matrix from the list (the 2nd element)
```

```{r fig.height = 3, fig.width = 3}
data = c(5,6,6,8,9,11,11,14,17,17,19,20,21,21,22,23,24,32,40,43,49)
mean(data, trim = 0.25) #trim mean
IQR(data) #iqr
summary(data) #fivenum
```

```{r fig.height = 2, fig.width = 2}
df = data.frame(data)
```
##Boxplot
```{r fig.height = 2, fig.width = 2}
ggplot(data = df, mapping = aes(y = data)) +
  geom_boxplot()
```
##Histogram
```{r fig.height = 2, fig.width = 2}
ggplot(data = df, mapping = aes(x = data)) +
  geom_histogram(bins = 10)
```
##QQ Plot
```{r fig.height = 2, fig.width = 2}
ggplot(data = df, aes(sample = data)) +
  geom_qq() + geom_qq_line()
```
##Scatter and Line (Ex. Flights)
```{r fig.height = 2, fig.width = 2}
data(flights)       #load data frame into session
#filter
all_alaska_flights <- flights %>%   #pipe operator
  filter(carrier == "AS")           #filter([expr related to colname])
#scatter
ggplot(data = all_alaska_flights, mapping = aes(x = dep_delay, y = arr_delay)) + 
    geom_point() # or geom_point(alpha = 0.2) or geom_jitter(width = 30, height = 30)
#line
early_january_weather <- weather %>% 
  filter(origin == "EWR" & month == 1 & day <= 15)
ggplot(data = early_january_weather, mapping = aes(x = time_hour, y = temp)) +
  geom_line()
#facets
ggplot(data = weather, mapping = aes(x = temp, fill = factor(month))) +
  geom_histogram(binwidth = 10, color = "white") + facet_wrap(~ month)
```

# Hypothesis Testing

Do men consume more hotwings than women?

Define the null and alternative hypothesis:

$H_0: \mu_M - \mu_F = 0$

$H_a: \mu_M - \mu_F > 0$

pvalue = 0.0008 < 0.05, so we reject $H_0$

There is enough evidence to conclude that, on average, men consume more hot wings than women

*Suppose that alpha = 0.000001*

pvalue = 0.0008 > alpha, so we fail to reject $H_0$

There is not enough evidence to conclude that, on average, men consume more hot wings than women

Load General Social Survery data into R (from section 1.7)

```{r}
library(resampledata)
data(GSS2002)
```

Create a table and a bar chart of the responses to the question about the death penalty.

```{r}
table(GSS2002$DeathPenalty)
ggplot(data = GSS2002, aes(x = DeathPenalty)) +
  geom_bar(fill = "red4")
```

Use the `table` command and the `summary` command in R on the gun ownership variable. What additional infomation does the `summary` command give that the `table` command does not?

```{r}
table(GSS2002$OwnGun)
summary(GSS2002$OwnGun)
```

Create a contingency table displaying the realtionship between opinions about the death penalty to that about gun ownership.

```{r}
table(GSS2002$DeathPenalty, GSS2002$OwnGun)
```

What proportion of gun owners favor the death penalty? Does it appear to be different from the proportion among those who do not own guns?

```{r}
gunprop <- GSS2002 %>%
  group_by(OwnGun) %>%
  summarize(prop = mean(DeathPenalty == "Favor", na.rm = TRUE))

kable(gunprop)
```