---
title: "HW 3, CH 6 MD HW"
author: "Matthew Swanson"
date: '20191022'
output: bookdown::html_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Orange Juice

```{r}
load("OJUICE.Rdata")
head(OJUICE)
```

1. Suppose a manufacturer wants to use to predict the sweetness ($y$) from the amount of pectin ($x$)
    
    i. Conduct Exploratory data analysis (Note: This constists of three parts)

a. Looking at the raw values:

```{r}
head(OJUICE)
#View(OJUICE)
```

b. Computing and describing the summary statistics likes means, medians, and standard deviations. Specially the correlation coefficient:

```{r}
library(skimr)
OJUICE %>% skim()

cor(x = OJUICE$PECTIN, y = OJUICE$SWEET)
```

* *The mean amount of pectin is 257 while the median is 244, meaning there are some high-value outliers*
* *The mean amount of sweetness is 5.66 while the median is 5.7. Sweetness follows a uniform distribution*
* *The std dev of pectin is 50 while the st dev of sweetness is 0.24*
* *75% of observed pectin values fall between 210 and 265*
* *The highest observed pectin value is 410*
* *25% of observed pectin values fall between 226 and 244*
* *Only 25% of observed pectin values are between 265 and 410*
* *The maximum sweetness is 5.2 while the maximum sweetness is 6*
* *50% of observed sweetness values fall between 5.5 and 5.82*
* *The correlation coefficient between pectin and sweetness is -0.48, meaning there is some negative correlation.*

c. Creating and describing data visualizations:

```{r}
library(ggplot2)
library(cowplot)

pp = ggplot(data = OJUICE, mapping = aes(x = PECTIN)) +
  geom_histogram(binwidth = 25, fill = 'red2', color = 'red4')
ps = ggplot(data = OJUICE, mapping = aes(x = SWEET)) +
  geom_histogram(binwidth = 0.1, fill = 'blue2', color = 'blue4')
plot_grid(pp, ps)

ggplot(data = OJUICE, mapping = aes(x = PECTIN, y = SWEET)) + 
    geom_point() +
    labs(x = "Pectin", y = "Sweetness", title = "Relationship Between Sweetness and Pectin") +
    geom_smooth(method = "lm")
```

* Most of the pectin observed values vall in between 215 and 285
* The sweetness values are pretty normally distributed, with slightly more falling between 5.7 and 6
* There is some negative correlation between sweetness and pectin

ii. Fit the model and write the equation.

``` {r}
sweet_model <- lm(SWEET ~ PECTIN, data = OJUICE) #y ~ x
sweet_model
```

> $\hat{Sweetness} = 6.252068 - 0.002311 \cdot Pectin$

iii. Give a practical interpretation of the value of b0, if possible.

> $b_0$ *is the y-intercept for our linear model. It could mean that for a pectin value of 0, you could expect a sweetness value of 6.25. However, since out lowest pectin value is around 200, that would be extrapolation, rendering* $b_0$ *as meaningless*

iv. Give a practical interpretation of the value of b1, if possible.

> *Every time we increase the pectin value by 1, our expected sweetness value decreases by -0.002311*

v. Predict the sweetness index if amount of pectin in the orange juice is 300 ppm.

`> 6.252068 - 0.002311 * 300`

    5.558768

vi. Figure out a way to do v. in R. (We did not discussed this in class)

```{r}
newx = data.frame(PECTIN = 300)
predicted_sweet = predict(sweet_model, newx)
predicted_sweet
```

vii. Conduct diagnostics and comment on how “good” the model is.

```{r}
library(moderndive)
regression_points <- get_regression_points(sweet_model)
regression_points

ggplot(sweet_model, aes(x = .fitted, y = .resid)) + geom_point()
```

> Based on the residual plot, it seems that the model is not very good. The correlation between sweetness and pectin is too low to create an accurate linear model.

# Credit

```{r}
library(ISLR)
data(Credit)
```

1. Suppose we want to use to predict the `Balance` ($y$) from the `Ethnicity` ($x$)

i. Conduct Exploratory data analysis

a. Looking at the raw values:

```{r}
head(Credit)
#View(Credit)
#Income in 10s of thousands
```

b. Computing and describing the summary statistics likes means, medians, and standard deviations. Specially the correlation coefficient:

```{r}
library(dplyr)
credit = Credit %>% select(Balance, Ethnicity)
credit %>% skim()

tbl = credit %>%
    group_by(Ethnicity) %>%
    summarize(Min = min(Balance), Mean = mean(Balance), Median = median(Balance), Max = max(Balance), SD = sd(Balance), N = n())
kable(tbl)

Balance_model <- lm(Balance ~ Ethnicity, data = credit)
Balance_model
summary(Balance_model)
```

* The minimum balance for all ethnicities was 0.
* Every ethnicity has a balance distribution that is heavily skewed towards a higher balance.
* AfricanAmericans have the highest mean balance at 531, Caucasians at 518, and Asians with the lowest at 512.
* Asians have the lowest median balance at 414, Caucasians at 465, and AfricanAmericans with the highest at 480.
* A Caucasian person holds by far the largest balance with 1999
* African Americans have the highest spread regarding balance ($\hat{\sigma}$ = 486)
* Caucasians have by far the lowest spread regarding balance ($\hat{\sigma}$ = 437)

c. Creating and describing data visualizations, individual and comparisons:

```{r}
ggplot(data = credit, mapping = aes(x = Balance, fill = factor(Ethnicity))) +
    geom_histogram(binwidth = 250, color = 'white') + 
    labs(x = "Balance", y = "Count", title = "Balance by Ethnicity") + 
    facet_wrap(~ Ethnicity, nrow = 1)
```

* Caucasians make up half of all observations
* The highest observed balance by far (~$2000) is of a Caucasian person
* Asians have more observed values around zero than AfricanAmericans

ii. Fit the model and write the equation.

$\hat{Balance} = 531.00 - 18.69 \cdot 1_{Asia}(x) - 12.50 \cdot 1_{Cauc}(x)$

iii. Give a practical interpretations of all coefficients and fitted values.

* $b_0 = Intercept = 531$: the mean balance for AfricanAmericans (our baseline)
* $b_{Asia} = EthnicityAsian = -18.69$: the difference in balance between Asians relative to AfricanAmericans, or in other words, on average Asians had balance $\$-18.69$ less.
* $b_{Cauc} = EthnicityCaucasian = -12.50$: the difference in balance between Caucasians relative to AfricanAmericans, or in other words, on average Caucasians had balance $\$-12.50$ less.
