---
title: "HW 2"
author: "Matthew Swanson"
date: "20191003"
output: 
    bookdown::html_document2:
    highlight: textmate
    theme: yeti
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.align = "center", comment = NA, options(scipen=999))
```

```{r}
library(resampledata)
data(FlightDelays)
library(dplyr)
library(knitr)
```

*Exercise 3.8*

In the Flight Delays Case Study in Section 1.1:

a. Find the 25% trimmed mean of the delay times for United Airlines and American Airlines.
b. Conduct a two-sided test to see if the difference in trimmed means is statistically significant.

```{r}
#a.
ans <- FlightDelays %>% 
  group_by(Carrier) %>% 
  summarize(Tr.mean = mean(Delay, trim = 0.25), N = n())
ans
```

$H_0:\mu_U - \mu_A = 0$

$H_A: \mu_U - \mu_A \neq 0$

```{r}
#b.
md <- ans[2, 2] - ans[1, 2]
observed <- md$Tr.mean

set.seed(1)
sims <- 10^4 - 1 # Number of simulations (iterations)
ans <- numeric(sims)
for(i in 1:sims){
  index <- sample.int(n = 4029, size =1123, replace = FALSE)
  ans[i] <- mean(FlightDelays$Delay[index], trim = 0.25) - mean(FlightDelays$Delay[-index], trim = 0.25)
}
pvalue <- 2*(sum(ans >= observed) + 1)/(sims + 1)
pvalue
```

> pvalue = $0.0002 < 0.05$, so we reject $H_0$

> There is enough evidence to conclude that, on average, the mean delay times between United Airlines and American Airlines are *not* equal

*Exercise 3.9*

In the Flight Delays Case Study in Section 1.1,

a. Compute the proportion of times the flights in May and in June were delayed more than 20 min, and conduct a two-sided test of whether the difference between months is statistically significant.

$H_0:p_J - p_M = 0$

$H_A: p_J - p_M \neq 0$

```{r}
FlightDelays$Delayed20 = FlightDelays$Delay > 20

ans <- FlightDelays %>%
  group_by(Month) %>%
  summarize(prop = mean(Delayed20), N = n())
ans

md <- ans[2, 2] - ans[1, 2]
observed <- md$prop
set.seed(1)
sims <- 10^4 - 1 # Number of simulations (iterations)
ans <- numeric(sims)
for(i in 1:sims){
  index <- sample.int(n = 4029, size =2030, replace = FALSE)
  ans[i] <- mean(FlightDelays$Delayed20[index]) - mean(FlightDelays$Delayed20[-index])
}
pvalue <- 2*(sum(ans >= observed) + 1)/(sims + 1)
pvalue
```

> pvalue = $0.0168 < 0.05$, so we reject $H_0$

> There is enough evidence to conclude that, on average, the proportion of delay times over 30 minutes during May versus during June are *not* equal

b. Compute the variance of the flight delay times in May and June and then conduct a two-sided test of whether the ratio of variances is statistically significantly different from 1.

$H_0:\sigma_J / \sigma_M = 1$

$H_A: \sigma_J / \sigma_M \neq 1$

```{r}
ans <- FlightDelays %>%
  group_by(Month) %>%
  summarize(SD = sd(Delay), N = n())
ans

md <- ans[2, 2] / ans[1, 2]
observed <- md$SD
set.seed(1)
sims <- 10^4 - 1 # Number of simulations (iterations)
ans <- numeric(sims)
for(i in 1:sims){
  index <- sample.int(n = 4029, size =2030, replace = FALSE)
  ans[i] <- sd(FlightDelays$Delay[index]) / sd(FlightDelays$Delay[-index])
}
pvalue <- 2*(sum(ans >= observed) + 1)/(sims + 1)
pvalue
```

> pvalue = $0.0336 < 0.05$, so we reject $H_0$

> There is enough evidence to conclude that, on average, the ratio of variances between flight delay times in May versus June are *not* equal